---
title: "Practical Session: working with jsonlite"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# https://hendrikvanb.gitlab.io/2018/07/nested_data-json_to_tibble/
# https://www.rdocumentation.org/packages/jsonlite/versions/1.7.2
# https://robotwealth.com/how-to-wrangle-json-data-in-r-with-jsonlite-purr-and-dplyr/
```

```{r}

```

## Load necessary packages

```{r}
library(jsonlite)
```

<br/>

## Reading JSON data with `fromJSON()`

<br/>

### Simplification: *when JSON arrays get converted to other R classes*

Depending on how the JSON array is structured, fromJSON() automatically converts the array from a list into a specific R class.

<br/>

#### Array of primitives (strings, numbers, booleans or null) 

When the JSON data is in the form of an array of strings, numbers, booleans or nulls (i.e. *primitives*), `fromJSON` converts it into a **vector** by default

```{r}
json_data <- '["Francesco", "Gabriel", "Simon", "Lisa"]'
fromJSON(json_data)
```

<br/>


