---
title: "Jsonbasics"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r packages_install, eval=FALSE, include=FALSE}
install.packages("knitr")
install.packages("rmarkdown")
install.packages('jsonlite')
library(knitr)
library(rmarkdown)
library(jsonlite)
```
## JSON

JSON (JavaScript Object Notation) is a lightweight data-interchange format. It is easy for humans to read and write. It is easy for machines to parse and generate. It is based on a subset of the JavaScript Programming Language Standard. 

JSON is a text format that is completely language independent but uses conventions that are familiar to programmers of the C-family of languages, including C, C++, C#, Java, JavaScript, Perl, Python, and many others. These properties make JSON an ideal data-interchange language.

## Why use JSON?
Here are the important benefits/ pros of using JSON:

- Provide support for all browsers
- Easy to read and write
- Straightforward syntax
- You can natively parse in JavaScript using eval() function
- Easy to create and manipulate
- Supported by all major JavaScript frameworks
- Supported by most backend technologies
- JSON is recognized natively by JavaScript
- It allows you to transmit and serialize structured data using a network connection.
- You can use it with modern programming languages.

## History of JSON
Here are important landmarks that form the history of JSON:

- Douglas Crockford specified the JSON format in the early 2000s.
- The official website was launched in 2002.
- In December 2005, Yahoo! starts offering some of its web services in JSON.
- JSON became an ECMA international standard in 2013.
- The most updated JSON format standard was published in 2017.

## Structure
JSON is built on two structures:

- A collection of name/value pairs. In various languages, this is realized as an object, record, struct, dictionary, hash table, keyed list, or associative array.
- An ordered list of values. In most languages, this is realized as an array, vector, list, or sequence.

These are universal data structures. Virtually all modern programming languages support them in one form or another. It makes sense that a data format that is interchangeable with programming languages also be based on these structures.
In JSON, they have the following rules:


## Rules for JSON Syntax

Rules for JSON syntax are:

- Data should be in name/value pairs
- Data should be separated by commas
- Curly braces should hold objects
- Square brackets hold arrays

## Data types in JSON 
|    Data Type   |Description | 
|:------|:-----|
| Number | It includes real number, integer or a floating number. |
| String | It consists of any text double-quoted with backslash escapement. |
| Boolean | The Boolean data type represents either True or False values. |
| Null | It denotes that the associated variable doesn’t have any value. |
| Object | It is a collection of key-value pairs and always separated by a comma and enclosed in curly brackets. |
| Array | It is an ordered sequence of values separated. |

## Number:
- The number is a double-precision floating-point format which depends on its implementation method.
- In JSON you can’t use Hexadecimal and Octal formats.

|  Type |Description | 
|:------|:-----|
| Integer | Number 1-9, and 0. Both positive and negative numbers. |
| Fraction | Fractions like 3. |
| Exponent | Exponent like e, e+. |

Example:

json-object-name = { string : number_value}
obj = {salary: 2600}


## String
Series of double-quoted Unicode characters with backslash escaping.

|  Type |Description | 
|:------|:-----|
| * | Use for double quotation typing. |
| / | Use for solidus. |
| \ | Use for reverse solidus. |
| B | Use to add backspace. |
| F | From feed. |
| N | To create a new line. |
| R | Use for carriage return. |
| T | To show horizontal tab. |
| U | Hexadecimal digits. |

## 

|  Type |Description | 
|:------|:-----|
| T | To show horizontal tab. |
| U | Hexadecimal digits. |



## Array

- It is an ordered collection of values.
- You should use an array when the key names are sequential integers.
- It should be enclosed inside square brackets which should be separated by ‘,’ (comma)



## JSON Object

A **JSON Object** is an entity in JSON which is enclosed in curly brackets. It is written in the unordered set of name and value pairs in which the name should be followed by “:” (colon), and the name/value pairs need to be separated using “,” (comma). It can be used when key names are arbitrary strings.

Syntax and example:

{ string :  value, ….. }

{
"id": 110,
"language": "Python",
"price": 1900,
}




## Application of JSON

- Helps you to transfer data from a server
- Sample JSON file format helps in transmit and serialize all types of structured data.
- Allows you to perform asynchronous data calls without the need to do a page refresh
- Helps you to transmit data between a server and web applications.
- It is widely used for JavaScript-based application, which includes browser extension and websites.
- You can transmit data between the server and web application using JSON.
- We can use JSON with modern programming languages.
- It is used for writing JavaScript-based applications that include browser add-ons.


## What is JSON not?

- Sample JSON data file is not a document format.
- It is not a markup language.
- JSON does not provide a general serialization format.
- It is not recurring or cyclical structures.
- It is also not an invisible structure.

## Disadvantages of JSON

Here are few advantages of JSON:

- No namespace support, hence poor extensibility
- Limited development tools support
- No support for formal grammar definition

## Sources

Those are the main sources used for the project:

- https://www.guru99.com/json-tutorial-example.html
- https://www.json.org/json-en.html 
